### 2023-03-21 Analysis tests.Rmd
Update scraper + upload test analysis rmd
Updated the scraper to extract the address line. Additionally added code to copy the scraper into the site-packages or dist-packages directory of the machine.

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(tidygeocoder)
```

```{r}
df <- read_csv("funda_buy_17-03-2023.csv")
```

```{r}
model <- lm(data = df, price~room +living_area + house_age + house_type)
summary(model)
```

```{r}

df_new <- df %>% 
  top_n(10) %>% 
  mutate(adddreszip = paste0(address, ", Noord-Holland")) %>% 
  geocode(adddreszip, method = 'osm', lat = latitude , long = longitude)

```

```{r}
easypackages::packages ("sf", "sp", "spdep", "spatialreg", "GWmodel", "tmap", "mapview", "car", "RColorBrewer", 
                        "cowplot", "leafsync", "leaflet.extras2", "mapview", "tidyverse")

```

```{r}
library(sf)
df_sp <- df_new %>% drop_na()
x = st_as_sf(df_sp, coords = c( "longitude", "latitude"), crs = 4326)
x <- st_transform(x, crs = 28992)
st_write(x, "test.gpkg")

```

```{r}

qtm(x)
```
```{r}
greendata_sp <- as_Spatial(x)

#fit the adaptive kernel 
#find adaptive kernel using gaussian function
abw <- bw.gwr(price~room +living_area + house_age,
             approach = "AIC", #specified by CV for cross-validation approach or by AIC corrected (AICc), we used AIC 
             adaptive = TRUE,
             kernel="gaussian", #this can be different function e.g., bisquare,exponential, depend on the prior understanding or choice of the modeler
             data=greendata_sp) #give the sp data created earlier
```

```{r}
#fitting the model with gwr.basic function
a.gwr <- gwr.basic(price~room +living_area + house_age, #the equation
             adaptive = TRUE,
             kernel="gaussian", #indicate the Kernel again
             bw = 17, #give the optimal bandwidth we found in the last stage
             data=greendata_sp)

#print the model result
a.gwr
```
```{r}
multiGWR <- gwr.multiscale(price~room +living_area + house_age, 
                        data = greendata_sp,
                        adaptive = T, 
                        max.iterations = 10, #usually large number to make sure each variable converage to best bandwidth
                        criterion="CVR", #criterion for determining the convergence of the back-fitting procedure
                        kernel = "bisquare", #kernel can also be gaussian, exponential
                        bws0=c(10,10,10), #starting number of neighbors  
                        verbose = F, predictor.centered=rep(T, 6)) 
```

```{r}
multiGWR
```

```{r}
library(leaflet)
leaflet(df_new) %>% addTiles() %>%
  addCircleMarkers(lng = ~longitude, lat = ~latitude, 
             popup = ~address)
```