### 2023-03-21 README.Rmd

---
title: "Explaining Noord-Holland house prices"
author: "Yúri Grings, Thomas Nibbering, Jitse Ruurd Nauta"
date: "2023-03-21"
output: github_document
---

```{=html}
<style type="text/css">
  body{
  font-size: 12pt;
  font-family:Cambria;
  text-align: justify;
}
headers{
text-align: left;

}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# "Explaining Noord-Holland house prices"

A Spatial Statistics and Spatial Machine Learning approach.

Included data of the research was scraped from Funda. Below map gives an overview of the scope of the data.

```{r}
library(leaflet)
df <- read.csv('data/test/funda_buy_21-03-2023(sample100p)latlon.csv')

leaflet(df) %>% addTiles() %>%
  addCircleMarkers(lng = ~longitude, lat = ~latitude, 
             popup = ~address)
```