### 2023-03-30 README.Rmd
Start with paper layout

---
title: "Explaining Noord-Holland house prices"
author: "Yúri Grings, Thomas Nibbering, Jitse Ruurd Nauta"
date: "2023-03-21"

output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)

```

A Spatial Statistics and Spatial Machine Learning approach.

Below the structure, results, and some code of our research can be found.

Kind regards,

Yúri Grings [Linkedin](https://www.linkedin.com/in/yurigrings/), Thomas Nibbering [Linkedin](https://www.linkedin.com/in/thomas-nibbering/), Jitse Ruurd Nauta [Linkedin](https://www.linkedin.com/in/jitseruurdnauta/)

## We used the Funda-Scraper package to scrape our data

The query below was used to scrape our final research data on (28-03-2023).

```{python eval = FALSE}
scraper = FundaScraper(area="provincie-noord-holland", want_to="buy", find_past=False,\
n_pages = 1000)
df = scraper.run()
```
Note that this is not the original Funda-Scraper code underneath. The Function is adjusted to contain additional adress information, compared to the origional function. Code is included in the Scraper.ipynb file to copy the updated package into the site-packages directory of python.

## A pre-processing pipeline is used to enrich data

OSM was used to enrich the scraped data. We geocode adressess into geometries objects for in order to make use of the geospatial models.
Additional spatial data includes:

* Distance to public transport (Bus, Subway, and Train)
* Distance to public services (Schools, Universities, Shopping centres)

## This enriched data is used to fit an OLS model

### OLS on scraped data

```{r eval = FALSE}
model <- lm(price~ room + bedroom + bathroom + living_area + house_age, data = funda_data)
```

### OLS on enriched data

```{r eval = FALSE}
model <- lm(price~ room + bedroom + bathroom + living_area + house_age + bus_dist +subway_dist + train_dist + university_dist + school_dist + mall_dist + supermarket_dist, data = funda_data)
```

### GWR model to adjust for spatial component in the data

To look at local spatial dependence, a GWR model was fit on the data.

```{python eval = FALSE}
#Run basic GWR in parallel mode

gwr_selector = Sel_BW(b_coords, b_y, b_X)
gwr_bw = gwr_selector.search(pool = pool)
print(gwr_bw)
gwr_results = GWR(b_coords, b_y, b_X, gwr_bw).fit(pool = pool)

```

### MGWR model to adjust for spatial component in the data

As it would be more realistic to have different band-widths for every feature, MGWR was used

```{python eval = FALSE}
#run MGWR in parallel mode. Note: max_iter_multi needs to be specified

mgwr_selector = Sel_BW(b_coords, b_y, b_X, multi=True)
mgwr_bw = mgwr_selector.search(pool=pool, max_iter_multi=200, criterion = "AICc") 
print(mgwr_bw)
mgwr_results = MGWR(b_coords, b_y, b_X, selector=mgwr_selector).fit(pool=pool)

```