### 2023-04-07 Paper/paper.Rmd
update

---
title: "Estimating the Effect of Urban Green Spaces on Residential House Prices in Amsterdam"
subtitle: "An Empirical Study using Remote Sensing and Geographically Weighted Regression" 
author: 
  - Yúri Grings
  - Thomas Nibbering
  - Jitse Ruurd Nauta
date: "2023-03-21"

output: 
    pdf_document:
        extra_dependencies: ["float", "booktabs"]
        latex_engine: xelatex
        pandoc_args: [
      "-V", "classoption=twocolumn"
    ]

subparagraph: yes
header-includes: |
  \usepackage{titlesec}
  \titlespacing{\section}{0pt}{10pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
  \titlespacing{\subsection}{0pt}{4pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
  \titlespacing{\subsubsection}{0pt}{5pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
  \usepackage{geometry}
  \geometry{left=1.5cm,right=1.5cm}
mainfont: Cambria
bibliography: references.bib
csl: apa.csl

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)

```

```{r include=FALSE}
#all code for plots
library(mapview)
library(tidyverse)
library(tmap)
library(sf)
setwd("./..")
funda_data <- st_read("data/Houseprices/funda_buy_amsterdam_31-03-2023_full_distances.gpkg")
PC4 <- st_transform(st_read("data/Amsterdam/PC4.json"), crs = 28992)
setwd("Paper")
```

# 1. Introduction
Green spaces play an important role in urban ecosystems. As a natural purifier for cities, these areas help control soil erosion [1], improve air quality [2, 3], reduce the urban heat island effect [4–6] and are considered to be aesthetically pleasing [7]. Moreover, green areas in cities are shown to provide notable improvements in residents’ physical and mental wellbeing [8, 9]. As the benefits of greenness in urban areas are increasingly recognised, numerous cities have developed urban greening programs to expand parks, build green roofs or plant more trees [10–12]. However, evaluating the effects of these initiatives to justify their costs remains a challenge, partly due to data sparsity.

A review of literature revealed that there are several studies that aimed to examine the effect of urban green spaces on residential house prices [13–19]. Often these studies adopted a hedonic pricing approach to estimate the economic value that individuals place on particular property characteristics, as reflected in the real-estate prices [14, 15]. However, most literature only focused on the distance to urban green areas as a valuation indicator [20, 21]. Here, a few studies considered a more comprehensive appreciation of greenness in cities by employing vegetation indices and the interaction with property and public amenities [16, 17, 22, 23].

Among the literature that explored the property value premium of urban green spaces more exhaustive, most adopted the normalised difference vegetation index (NDVI) [16, 22, 23]. Here, the relative abundance of vegetation was measured across varying buffer distances using remote sensing systems with various spatial resolutions. Moreover, most of these studies also considered the interaction with residential and public amenities, such as the number of rooms or proximity to convenience stores [16, 23]. However, these analyses were mainly performed in cities with conservative urban greening programs in place [23]. Lastly, the adoption of NDVI for identifying (urban) green spaces is often criticised due to its sensitivity to varying spatial scales and inability to differentiate between vegetation types [24]. Nonetheless, the index is widely accepted and can be considered almost a standard approach [25, 26].

Based on the examined literature, a hedonic pricing approach in combination with the normalised difference vegetation index and the interaction with particular amenities emerges as an effective method to estimate the effect of urban green spaces on residential house prices. Given the limited literature available on property value premiums associated with greenness exposure in cities with more progressive urban greening programs, this study aims to answer the research question of what the economic value of urban green spaces in the city of Amsterdam is in the year 2023, as reflected in the real-estate prices.

# 2. Methods
## 2.1 Study Area
The geographical scope in this study concentrated on the city of Amsterdam, as presented in Figure 1. The city is situated in the Western part of the Netherlands and home to almost 880,000 residents [27] that occupy a space of 219.3 km2  [28]. As a result, the city is considered one of the most densely populated in the country. Surrounding the residential areas, the vast majority of land is occupied with vegetation [29]. Moreover, the municipality has progressive urban greening programs in place, stimulating and obligating residents and city districts to participate in greening initiatives [12]. As a result, these policies aimed at expanding urban green spaces have the potential to increase nearby housing premiums.

```{=latex}
\begin{figure}[t]
  \centering
  \includegraphics[width=\linewidth]{figures/study_area.png}
  \caption{Overview of the study area.}
\end{figure}
```

## 2.2 Data
In this study, property listings were obtained from Funda.nl, a Dutch online real-estate platform [30]. Here, data included the sale price and property characteristics of residential houses in Amsterdam listed on the platform on 31 March 2023. Moreover, information about multiple public amenities in Amsterdam were obtained from a variety of sources, including ArcGIS [31], OpenStreetMap.org [32, 33] and the municipality of Amsterdam [34]. Selection of public amenities, used as controlling variables, were based on applicable literature, as presented in Table 1. Adjusting for these variables was done to limit the effect of potential confounders on the association of interest. Finally, remote sensing images in red and near-infrared were obtained from Sentinel-2 in August 2022 [35]. This allowed on the capitalisation of high greenness levels in that period, leading to improved differentiation between green and urban areas.

## 2.3 Data Enrichment

To examine the economic value of urban green spaces as reflected in the real-estate prices, coordinates for each property were obtained using geocoding and transformed into a single coordinate reference system, i.e. EPSG 28992, to ensure consistency and accuracy of the analysis. Exclusion criteria were in place for properties for which coordinates could not be obtained, i.e. future housing projects. Moreover, straight line distances from each property to the nearest public amenities were enumerated using the nearest neighbour algorithm in combination with Euclidean distance. This is a common approach to adjust for the implicit premium of local public amenities on house prices.

Furthermore, urban green vegetation was measured using NDVI. This vegetation index was derived from the Sentinel-2 satellite, as expressed in Equation 1. Here, N and R refer to the spectral reflectance measurements obtained in the red and near-infrared regions, respectively. In general, the index ranges from -1.0 to 1.0 with positive values indicating greenness. The usage of this remote sensing system provided more accurate images for identifying green spaces, i.e. 10x10m, compared to the moderate resolution of Landsat, i.e. 30x30m.

Surrounding the residential properties, mean values of NDVI were obtained at varying spatial scales, i.e. 100, 300 and 500m. In doing so, the variation of greenness around the property was included in the analysis. Moreover, the usage of these particular thresholds ensured that the results were less vulnerable to scaling effects given the resolution of the input data. Finally, based on the extraction and enhancement of data, several independent variables were identified and constructed, as presented in Table 1. These variables were included in the hedonic pricing model, as will be elaborated on in Section 2.4, to either adjust or explain the economic value of urban green areas in Amsterdam.

## 2.4 Analysis

```{=latex}
\begin{table*}[b]
\centering
\caption{Descriptive Statistics of Property, Public Amenity and Greenness Characteristics of Properties in  Amsterdam.}
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Independent Variable}      & \textbf{Mean} & \textbf{SD} & \textbf{Description}                        \\ \midrule
\multicolumn{4}{c}{\textbf{Summary Statistics (N = 1,719)}}                                                    \\ \midrule
\textbf{Property Characteristics}  &               &             &                                             \\
Living Area (m2)                   & 105.7         & 71.0        & Size of livable area of the property        \\
Number of Bedrooms                 & 2.5           & 1.4         & Number of bedrooms of the property          \\
Number of Bathrooms                & 1.1           & 0.7         & Number of bathrooms of the property         \\
House Age (years)                  & 76.9          & 63.2        & Age of the property in 2023                 \\
\textbf{Design}                    &               &             &                                             \\
NDVI100                            & 0.200         & 0.073       & Average NDVI at 100 m around property       \\
NDVI300                            & 0.212         & 0.069       & Average NDVI at 300 m around property       \\
NDVI500                            & 0.216         & 0.067       & Average NDVI at 500 m around property       \\
\textbf{Transit Distance}          &               &             &                                             \\
Distance to Train Station (m)      & 1703.7        & 958.2       & Euclidean distance to nearest train station \\
Distance to Tram Station (m)       & 689.1         & 1024.7      & Euclidean distance to nearest tram station  \\
Distance to Metro Station (m)      & 1303.2        & 951.8       & Euclidean distance to nearest metro station \\
\textbf{Destination Accessibility} &               &             &                                             \\
Distance to City Center (m)        & 3770.5        & 2288.1      & Euclidean distance to city center           \\
Distance to Business District (m) & 4659.2 & 2228.3 & Euclidean distance to business district \\
Distance to School (m)             & 268.3         & 183.8       & Euclidean distance to nearest school        \\
Distance to Convenience Stores (m) & 266.7         & 206.9       & Euclidean distance to nearest store         \\ \bottomrule
\end{tabular}
\end{table*}
```

## 2.4.1 Ordinary Least Squares

## 2.4.2 Spatial Dependence and Heterogeineity

## 2.4.3 Geographically Weighted Regression

# 3. Results and Discussion

```{=latex}
\begin{figure*}[ht]
  \centering
  \includegraphics[width=\linewidth]{figures/ndvi_houseprices.png}
  \caption{This is a figure.}
\end{figure*}
```

# 4. Conclusion

# 5. References